\chapter{Introduction}

\section{Context}

The healthcare sector is undergoing a significant transformation driven the Internet of Things (IoT) technologies.
Continuous, real-time monitoring of patient vital signs and ambient room conditions is a foundational requirement in any clinical environment, enabling medical staff to detect deterioration early, respond rapidly, and reduce the administrative overhead associated with manual observation rounds.

This project proposes a complete, end-to-end hospital monitoring solution composed
of four integrated subsystems:

\begin{itemize}
	\item \textbf{Embedded acquisition node (ESP32):} real-time sensor acquisition and authenticated cloud transmission.
	\item \textbf{Backend server:} database management, alert rule evaluation, and
	      user authentication.
	\item \textbf{Web dashboard:} real-time data visualisation, historical charts,
	      and alert management for medical staff.
	\item \textbf{AI prediction layer:} anomaly detection on patient vitals and
	      environmental condition assessment using cloud-hosted machine learning models.
\end{itemize}

\section{Problem Statement}

Hospital rooms require continuous monitoring of two categories of parameters:
environmental conditions (temperature and humidity), which directly affect patient
comfort, infection risk, and equipment operation; and patient vital signs (heart rate
and blood oxygen saturation), which are primary indicators of physiological status.

The current approach in many low-resource clinical settings relies on periodic manual
measurement, which introduces observation gaps and increases nursing workload.

\section{Objectives}

The project objectives are divided into functional and technical categories.

\subsection*{Functional Objectives}

\begin{enumerate}
	\item Continuously acquire temperature (°C), humidity (\%), heart rate (BPM), and
	      blood oxygen saturation (SpO\textsubscript{2}, \%) from dedicated sensors.
	\item Transmit sensor data to a cloud database at a near-real-time rate of up to 1\,Hz.
	\item Implement a web dashboard for real-time data visualisation, history browsing,
	      and alert management.
	\item Generate automated predictions and anomaly alerts via a cloud AI model.
\end{enumerate}

\subsection*{Technical Objectives}

\begin{enumerate}
	\item Availability: the system shall operate continuously with network recovery mech-
	      anisms.
	\item Reliability: invalid or inconsistent data shall be detected and flagged.
	\item Security: secure authentication and encrypted communication (HTTPS).
	\item Performance: near real-time data updates based on sampling frequency.
\end{enumerate}
