% ============================================================
%  CHAPTER 7 — Discussion
% ============================================================

\chapter{Discussion}
\label{ch:discussion}

% ============================================================
\section{Validity of Results}
% ============================================================

\subsection{Embedded Subsystem}

The sensor measurements are consistent with the manufacturer-specified accuracy of the
DHT22 ($\pm$0.5\,°C, $\pm$2--5\,\% RH) and within the expected operational range of
the MAX30102 for resting-state adult subjects.
The BPM standard deviation of 2.3\,BPM reflects the inherent variability of
beat-to-beat interval detection under minor motion artefacts.
The SpO\textsubscript{2} accuracy of $\pm$1\,\% is within the $\pm$2\,\% tolerance
generally accepted.

\subsection{Backend and AI Subsystems}
The backend results demonstrate high consistency in procedure execution, with tRPC overhead remaining negligible compared to the Firebase round-trip time. The AI forecasting accuracy (MAE of 1.2 BPM) is clinically valid for providing early-warning signals, although its performance is highly dependent on the quality of the incoming 1Hz telemetry stream. Anomaly detection showed high precision (0.94), successfully filtering out transient sensor noise while capturing sustained physiological deviations.

\subsection{Web Dashboard Subsystem}
The dashboard's performance is stable under continuous operation, with the WebWorker-based face recognition maintaining accurate personnel logs without impacting the real-time chart frame rates. The responsive design ensures that clinicians can monitor vitals across different device form factors with no loss of data fidelity.

% ============================================================
\section{Limitations}
% ============================================================

\subsection{Embedded Subsystem}

\begin{description}
	\item[No local offline buffer.]
	      Records generated during network outages are discarded.

	\item[Plaintext NVS storage.]
	      Firebase credentials (API key, email, password) are stored in NVS without
	      encryption.
	      The ESP32 NVS partition supports native encryption via an NVS key partition;
	      this should be enabled before any production deployment.

	\item[Implicit shared-memory atomicity.]
	      The sensor globals are 32-bit floats written on Core~1 and read on Core~0
	      without explicit synchronisation.
	      A FreeRTOS mutex or atomic qualifier is the correct long-term solution.

	\item[Single patient per device.]
	      The current design maps one device to exactly one patient and one room.
	      Multi-patient rooms would require either multiple devices or a structural change
	      to the data model.
\end{description}

\subsection{Backend Subsystem}
\begin{description}
	\item[No rate limiting.] The current tRPC implementation does not enforce rate-limiting, which could make the system vulnerable to denial-of-service attacks if exposed to the public internet.
	\item[SMTP delivery latency.] Email alerts are dependent on external SMTP relay availability, which may introduce latencies of 2--5 seconds during network congestion.
	\item[Single-region deployment.] The backend and Firebase services are currently deployed in a single geographic region, increasing latency for cross-regional users and lacking multi-region failover.
\end{description}

\subsection{Web Dashboard Subsystem}
\begin{description}
	\item[High resource consumption.] The real-time chart rendering and local face-recognition models impose significant CPU and GPU load on the browser, which may reduce battery life on mobile devices.
	\item[Browser-specific optimizations.] While cross-browser compatible, the system performs best in Chromium-based browsers due to advanced WebGL and WebWorker optimizations.
	\item[No persistent offline mode.] While the dashboard handles transient disconnections, it does not support full offline clinical record access.
\end{description}

\subsection{AI Subsystem}
\begin{description}
	\item[Linear trend assumptions.] The current forecasting model uses linear regression, which may fail to capture complex non-linear physiological events such as sudden cardiac arrythmias.
	\item[Small baseline window.] The 20-point analysis window is sufficient for short-term insights but lacks the context of multi-hour or multi-day historical trends.
	\item[Data quality dependency.] The accuracy of anomaly detection is directly coupled to the ESP32's signal filtering; poor sensor contact results in a higher false-positive rate.
\end{description}

% ============================================================
\section{ESP32 Constraints Impact}
% ============================================================

The most significant platform constraint encountered was the non-thread-safety of
the mbedTLS context, which imposed the non-obvious architectural requirement of
deferring Firebase initialisation until the task is running on Core~0.

The 8\,KiB task stack was found to be adequate for the TLS handshake and async send
operations, with an estimated 1.5\,KiB headroom.

The choice of string-based JSON construction introduces heap fragmentation over long
sessions.

% ============================================================
\section{Robustness}
% ============================================================

The unified \texttt{wifiConnect()} function using \texttt{WiFi.begin()} handles both
cold-start failure and mid-session dropout correctly, as confirmed by Session~2 and
Session~3 of the stability tests.

The delta-threshold filter provides robustness against spurious sensor noise:
fluctuations smaller than 0.3\,°C, 0.5\,\% RH, or 2\,\% SpO\textsubscript{2} do
not generate database writes, reducing both network load and the volume of
low-information records in the database.

% ============================================================
\section{Identified Improvements}
% ============================================================

Priority improvements for future versions of the embedded subsystem:

\begin{enumerate}
	\item \textbf{Local offline buffer.}
	      A LittleFS-backed FIFO queue for records generated during network outages,
	      with automatic retry on reconnection.
	      This would eliminate the record loss observed in Sessions~2 and~3.

	\item \textbf{NVS encryption.}
	      Enable the ESP32 native NVS encryption partition to protect credentials at rest.

	\item \textbf{Thread-safe sensor globals.}
	      Replace raw float globals with a mutex-protected struct or a FreeRTOS queue
	      for safe inter-core communication.

	\item \textbf{Wi-Fi modem sleep.}
	      Implement modem sleep between upload intervals to reduce average current usage for future battery-powered deployments.
\end{enumerate}

\begin{enumerate}
    \item \textbf{Deep Learning Models.} Migrating the AI layer to an LSTM (Long Short-Term Memory) or Transformer architecture to better capture non-linear physiological trends.
    \item \textbf{PWA Support.} Implementing Progressive Web App (PWA) features to enable offline data caching and native push notifications for critical clinical alerts.
    \item \textbf{Distributed Cache.} Adding a Redis-based caching layer for the tRPC backend to accelerate historical data retrieval for long-term patient records.
    \item \textbf{Multi-Factor Authentication.} Enhancing security for clinical staff logins using hardware security keys or biometric verification.
\end{enumerate}
