%  CHAPTER 6 — Experimental Results and Validation

\chapter{Experimental Results and Validation}
\label{ch:results}

\section{Experimental Protocol}

\subsection{Embedded Subsystem Protocol}

\begin{description}
	\item[Hardware setup.]
	      One ESP32-WROOM-32 development board, one DHT22 sensor, one MAX30102 breakout
	      board, all powered via USB from a laptop running the serial monitor.

	\item[Network environment.]
	      2.4\,GHz 802.11n Wi-Fi network.
	      Firebase Realtime Database.

	\item[Firmware build.]
	      Debug build (\texttt{deploy.sh --code --debug}) for all experiments, to capture
	      serial output.
\end{description}

\subsection{Backend Protocol}
The backend was validated using a \textbf{tRPC server-side caller} and \textbf{Vitest} for unit testing. The test environment consisted of a local Node.js v20 runtime and a Firestore emulator. A total of 15 integration test cases were executed, covering role-based authentication, schema validation, and SMTP relay reliability.

\subsection{Web Dashboard Protocol}
The dashboard was tested across Chrome and Firefox on both Windows and Linux. Real-time update latency was measured using the \textbf{Chrome DevTools Performance profiler}. Test scenarios included multi-patient monitoring views (up to 5 concurrent streams) and automated face detection triggers for authorized and unauthorized personnel.

\subsection{AI Protocol}
The AI subsystem was evaluated using a \textbf{synthetic clinical dataset} of 500 data points containing simulated vital sign transitions (e.g., resting to tachycardia). Evaluation metrics included \textbf{Mean Squared Error (MSE)} for forecasting accuracy and \textbf{F1-score} for anomaly detection precision and recall. Tests were performed on the Node.js service layer with a 20-point sliding window.

\section{Software Results (PC / Server)}

\begin{table}[H]
	\centering
	\caption{Backend API performance metrics.}
	\label{tab:backend_results}
	\begin{tabular}{lll}
		\toprule
		\textbf{Endpoint}       & \textbf{Latency (Avg)} & \textbf{Success Rate} \\
		\midrule
		\texttt{people.list}    & 45\,ms                 & 100\%                 \\
		\texttt{patients.get}   & 32\,ms                 & 100\%                 \\
		\texttt{events.log}     & 28\,ms                 & 99.8\%                \\
		\texttt{ai.getInsights} & 55\,ms                 & 99.5\%                \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{AI model evaluation results.}
	\label{tab:ai_results}
	\begin{tabular}{lll}
		\toprule
		\textbf{Task}     & \textbf{Metric} & \textbf{Value} \\
		\midrule
		Anomaly Detection & F1-Score        & 0.92           \\
		Anomaly Detection & Precision       & 0.94           \\
		Vital Forecasting & MAE (HR)        & 1.2 BPM        \\
		Vital Forecasting & RMSE (SpO2)     & 0.8\%          \\
		\bottomrule
	\end{tabular}
\end{table}

% ============================================================
\section{Embedded Results (ESP32)}
% ============================================================

Table~\ref{tab:sensor_accuracy} reports the accuracy of the ESP32 sensor readings.

\begin{table}[H]
	\centering
	\caption{ESP32 sensor accuracy validation in an indoor environment).}
	\label{tab:sensor_accuracy}
	\begin{tabular}{lcccc}
		\toprule
		\textbf{Parameter}        & \textbf{Reference}        & \textbf{Mean Measured}
		                          & \textbf{Mean Abs.\ Error} & \textbf{Std.\ Dev.}                          \\
		\midrule
		Temperature (°C)          & 26.0\,°C                  & 26.1\,°C               & 0.1\,°C  & 0.15\,°C \\
		Humidity (\%)             & 53.0\,\%                  & 53.4\,\%               & 0.4\,\%  & 0.8\,\%  \\
		Heart Rate (BPM)          & 72\,BPM                   & 73\,BPM                & 1.0\,BPM & 2.3\,BPM \\
		SpO\textsubscript{2} (\%) & 98\,\%                    & 97\,\%                 & 1.0\,\%  & 0.9\,\%  \\
		\bottomrule
	\end{tabular}
\end{table}

Table~\ref{tab:stability} reports the outcome of the three operational sessions,
focusing on Wi-Fi recovery and data continuity.

\begin{table}[H]
	\centering
	\caption{System stability and recovery validation results.}
	\label{tab:stability}
	\begin{tabular}{lcc}
		\toprule
		\textbf{Scenario}               & \textbf{Recovery Time} & \textbf{Records Lost}             \\
		\midrule
		Normal 15-min session           & N/A                    & 0                                 \\
		Mid-session Wi-Fi dropout       & $\approx$12\,s         & 12 (1 per second during outage)   \\
		Cold-start, Wi-Fi delayed 30\,s & 30\,s                  & 30 (not retried; no local buffer) \\
		\bottomrule
	\end{tabular}
\end{table}

% ============================================================
\section{Latency and FPS Analysis}
% ============================================================


\begin{table}[H]
	\centering
	\caption{System-wide latency and FPS comparison.}
	\label{tab:latency_full}
	\begin{tabular}{lccl}
		\toprule
		\textbf{Stage}              & \textbf{PC/Server}               & \textbf{ESP32}      & \textbf{Analysis}           \\
		\midrule
		Sensor acquisition          & N/A                              & $\leq$2\,000\,ms    & Hardware constraint (DHT22) \\
		Local processing            & N/A                              & $<$5\,ms            & Negligible                  \\
		HTTPS upload (steady state) & N/A                              & 120--280\,ms        & RTT to Firebase             \\
		Backend API response        & 28--55\,ms                       & N/A                 & tRPC overhead               \\
		AI inference                & 8--15\,ms                        & N/A                 & Vector processing           \\
		End-to-end (sensor to UI)   & \multicolumn{2}{c}{180--450\,ms} & Real-time compliant         \\
		\midrule
		\textbf{FPS — Dashboard UI} & \textbf{60\,FPS}                 & N/A                 & Recharts render budget      \\
		\textbf{FPS — Face detect.} & \textbf{$\approx$15\,FPS}        & N/A                 & WebWorker inference rate    \\
		\textbf{FPS — Sensor push}  & N/A                              & \textbf{1\,Hz}      & Upload interval constraint  \\
		\bottomrule
	\end{tabular}
\end{table}

The dashboard rendering pipeline achieves a sustained 60\,FPS for chart updates,
computed as $\text{FPS} = 1 / T_{\text{inference}} = 1 / 16.7\,\text{ms}$.
The face detection WebWorker, running the face-api.js SSD MobileNet model, achieves
approximately 15\,FPS ($T_{\text{inference}} \approx 67\,\text{ms}$ per frame) on a
modern desktop CPU, which is adequate for room-entry tracking.
The embedded sensor push rate is fixed at 1\,Hz by the \texttt{FIREBASE\_INTERVAL\_MS}
parameter, yielding an effective sensor FPS of 1\,frame/s for patient vitals and
0.5\,frame/s for room data (constrained by the DHT22 2\,s minimum inter-sample interval).

% ============================================================
\section{Comparison: Design Specification vs.\ Observed Behaviour}
% ============================================================

\subsection{Specification vs.\ Implementation (Embedded Subsystem)}

Table~\ref{tab:comparison} compares the functional requirements from the project
specification against the observed behaviour of the implemented embedded subsystem.

\begin{table}[H]
	\centering
	\caption{Specification vs.\ implementation validation for the embedded subsystem.}
	\label{tab:comparison}
	\begin{tabularx}{\linewidth}{lXX}
		\toprule
		\textbf{Requirement} & \textbf{Expected}                                             & \textbf{Observed} \\
		\midrule
		Wi-Fi reconnection
		                     & Automatic, no intervention
		                     & \cmark\ Reconnects within 10--15\,s                                               \\
		Firebase authentication
		                     & Email/password over TLS
		                     & \cmark\ JWT token obtained; events logged via callback                            \\
		Delta-threshold filtering
		                     & Suppress redundant writes
		                     & \cmark\ Confirmed via Firebase console (no duplicate records)                     \\
		AP configuration portal
		                     & Accessible at 192.168.4.1
		                     & \cmark\ Served from LittleFS; settings persisted to NVS                           \\
		NVS persistence
		                     & Settings survive reboot
		                     & \cmark\ Verified across 5 cold reboots                                            \\
		Debug/release build
		                     & Zero serial overhead in release
		                     & \cmark\ Confirmed via binary size reduction                                       \\
		\bottomrule
	\end{tabularx}
\end{table}

\begin{table}[H]
	\centering
	\caption{Specification vs.\ implementation validation for backend and web.}
	\label{tab:comparison_full}
	\begin{tabularx}{\linewidth}{lXX}
		\toprule
		\textbf{Requirement} & \textbf{Expected}          & \textbf{Observed}           \\
		\midrule
		Real-time Updates    & Update UI in < 1\,s        & \cmark\ Avg latency 320\,ms \\
		AI Forecasting       & 5-minute trend accuracy    & \cmark\ MAE < 2.0 BPM       \\
		Face Recognition     & Authorized staff login     & \cmark\ 92\% accuracy at 2m \\
		Email Alerting       & Dispatch on critical event & \cmark\ Delivered in < 5\,s \\
		\bottomrule
	\end{tabularx}
\end{table}

\subsection{Theoretical Model vs.\ Observed System Behaviour}

The latency model defined in Chapter~4 (Equation~\ref{eq:latency}) predicted
a steady-state end-to-end latency in the range 225--305\,ms for patient data
(excluding the DHT22 2\,s acquisition floor).
The observed end-to-end latency of 180--450\,ms is consistent with this model:
the lower bound matches low-contention Wi-Fi conditions where the TLS session is
already established, while the upper bound corresponds to periods of elevated
network RTT or transient Firebase congestion.

The delta-threshold filter was predicted to suppress the majority of redundant
uploads during steady-state monitoring (constant vitals).
In practice, during a 15-minute resting session, the filter reduced the total
number of Firebase write operations by approximately 60\,\% compared to an
unconditional 1\,Hz upload strategy, confirming the model's design intent.

The AI forecasting model, based on OLS linear regression over a 20-point window,
was designed with the assumption that short-term vital sign trends are
approximately linear.
This assumption holds well under stable resting conditions (MAE of 1.2\,BPM),
but degrades during rapid transitions such as post-exercise tachycardia, where
the linear model underestimates the rate of change.
This is consistent with the known limitation of the linear extrapolator and
motivates the future work on LSTM-based models.

% ============================================================
\section{Error Analysis}
% ============================================================

\subsection{Embedded Subsystem}

Three categories of error were identified and analysed:

\textbf{Sensor read errors.}
The DHT22 occasionally returns NaN on the first read after power-up due to its
required 1-second warm-up delay.
The NaN guard in \texttt{updateSensors()} discards the invalid reading silently.
No invalid values were written to the database in any test session.

\textbf{Records lost during outage (open).}
Records generated during a Wi-Fi outage are discarded.
The system does not implement local buffering or retry-on-reconnect.

\subsection{Backend, Web Dashboard, and AI Subsystems}

\textbf{Authentication Timeouts.}
During long monitoring sessions, Firebase ID tokens may expire. The dashboard was updated to include an \texttt{authInterceptor} that automatically refreshes the token, ensuring uninterrupted API access.

\textbf{Web Worker Contention.}
Running face-api.js on the main thread caused occasional UI stutter during chart rendering. This was resolved by moving the neural network inference to a dedicated Web Worker, isolating the computational load from the rendering thread.

\textbf{AI False Positives.}
The Z-score model occasionally flagged sensor "noise" (e.g., finger movement) as a clinical anomaly. To mitigate this, a temporal persistence filter was added: an anomaly is only reported if it persists for more than 3 consecutive samples.
